### 2019-05-13 doc/bilevel/paper.tex
Slight reorg.

@@

\SANDsection{Motivating Problems}
\label{sec:problems}

In multilevel optimization problems, a subset of decision variables
at each level is constrained to take values associated with an
optimal solution of a distinct, lower level optimization problem.
For example, a general formulation for bilevel programs is
\begin{equation}
\label{eqn:bilevel}
\begin{array}{ll}
\min_{x \in X, y}   & F(x,y)\\
\st                 & G(x,y) \leq 0\\
                    & y \in P(x)
\end{array}
\end{equation}
where
\[
\begin{array}{llll}
P(x) & = & \argmin_{y \in Y}    & f(x,y)\\
 & &                            & g(x,y) \leq 0
\end{array}
\]
$P(x)$ defines a \textit{lower-level problem}, which may have multiple solutions.
Here $x$ is the primary upper-level decision, and $y$ is the anticipated lower-level 
decision.

@@

>>>>

\if 0

% WEH - For now, let's omit this introduction since it needs to be 
%       reworked anyway.

@@

\fi

\SANDsection{Expressing Multilevel Programs}
\label{sec:modeling}

In multilevel optimization problems, a subset of decision variables
at each level is constrained to take values associated with an
optimal solution of a distinct, lower level optimization problem.
For example, a general formulation for bilevel programs is
\begin{equation}
\label{eqn:bilevel}
\begin{array}{ll}
\min_{x \in X, y \in Y}   & F(x,y)\\
\st                 & G(x,y) \leq 0\\
                    & y \in P(x)
\end{array}
\end{equation}
where
\[
\begin{array}{llll}
P(x) & = & \argmin_{y \in Y}    & f(x,y)\\
 & &                            & g(x,y) \leq 0
\end{array}
\]
$P(x)$ defines a \textit{lower-level problem}, which may have multiple solutions.
Here $x$ is the primary upper-level decision, and $y$ is the anticipated lower-level 
decision.

@@

\textbf{TODO: Describe pessimistic problems here.}

Many authors use a more compact form to express (\ref{eqn:bilevel}):
\begin{equation}
\label{eqn:bilevel-compact}
\begin{array}{lll}
\min_{x \in X, y \in Y}   & F(x,y) & \\
\st                       & G(x,y) \leq 0 & \\
                          & \min_{y \in Y}    & f(x,y)\\
                          &                   & g(x,y) \leq 0
\end{array}
\end{equation}
This notation defines an implied constraint by not explicitly referencing the solution to the lower-level problem.

The \code{pyomo.bilevel} package uses the \code{SubModel} component to declare a block that defines a lower-level problem. 
For example, consider:
\begin{qlisting}
model = ConcreteModel()
model.x = Var(bounds=(1,2))
model.v = Var(bounds=(1,2))
model.sub = SubModel()
model.sub.y = Var(bounds=(1,2))
model.sub.w = Var(bounds=(-1,1))

model.o = Objective(expr=model.x + model.sub.y + model.v)
model.c = Constraint(expr=model.x + model.v >= 1.5)
model.sub.o = Objective(expr=model.x+model.sub.w, sense=maximize)
model.sub.c = Constraint(expr=model.sub.y + model.sub.w <= 2.5)
\end{qlisting}
This Pyomo model defines four variables, \code{x}, \code{v}, \code{sub.y} and \code{sub.w}.
Variables \code{x} and \code{v} are declared in the upper-level
problem, and \code{v} only appears in the upper-level problem.
Variables \code{sub.y} and \code{sub.w} are declared in the submodel.
However, note that the \code{sub.y} variable appears in the upper-level
problem, while the \code{sub.w} variable only appears in the
lower-level problem.

While this syntax mimics the structure of (\ref{eqn:bilevel-compact}), it adds additional abstractions that make it difficult to interpret:
\begin{itemize}

\item It can be difficult to easily discern which are the upper and lower level variables.  In general, Pyomo does not place strong restrictions on the location of variable declarations relative to the location where those variables are used in expressions.  Thus, while a hierarchical block structure suggests nested model representations, this is not strongly enforced when objective and constraint expressions are generated.\\
\mbox{}\\
IDEA: Enforce a stronger semantics on PAO block representations for bilevel programs.\\
IDEA: Use model annotations to denote upper- and lower- level variables.\\

\item This syntax does not indicate how the solution of the lower level relates to constraints in the upper level problem.  Specifically, there is no explicit indication how solutions to the lower level problem relate to variables in the upper-level problem.  In simple applications, where there is a direct correspondence of variables (using the same names), this simple syntax may be sufficient.  But in more complex applications the variables need not have the same naming in the upper- and lower-level problems.  Thus, PAO needs a mechanism to associate these variables.

\end{itemize}

There are also practical issues that reflect the needs of users with
respect to the solution of bilevel programs.  Generally, it is assumed
that the solution of~(\ref{eqn:bilevel}) consists of the optimal upper-
and lower-level variables.  However in practice a user may only need
the optimal upper-level variables.  If true, this may allow for a
faster response.  Additionally, there's another solution response that
may be of interest to analysts, where solution consists of the optimal
upper-level variables along with other diagnostic information related
to the solution of the lower-level problem.  For example, in some cases
the dual of the lower-level problem can be used, so the solution to the
problem could include the dual lower-level variables.  Note that these
considerations also generalize to multi-level problems, where a user may
wish to specify the requirements for a solution value (e.g. needing only
the first- and second-level variables in a tri-level problem).

The following subsections describe ideas for expressing and solving multi-level programs, with the
goal of defining some alternatives that the PAO developers can discuss while refining the 
problem representations in PAO.

\subsection{TODO}

TODO: Discuss a new modeling idea here.

\subsection{TODO}

TODO: Discuss a new modeling idea here.

\SANDsection{Special Cases}
\label{sec:problems}
